---
title: "Homework 1 - Time Series Visualization"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2) 
```

# Exercise 1
Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

```{r, echo=TRUE, cache=TRUE, eval=FALSE}
help(gold)
help(woolyrnq)
help(gas)
# to do

```

Use `autoplot` to plot each of these in separate plots.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)
```

**What is the frequency of each commodity series?**
Hint: apply the `frequency()` function.

Gold: täglich
wool: quartal
gas: monatlich
```{r, echo=TRUE, cache=TRUE}
# to do
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

Use `which.max()` to spot the outlier in the `gold` series. 
**Which observation was it?**

```{r, echo=TRUE, cache=TRUE}
# to do
which.max(gold)
gold[which.max(gold)]

```

# Exercise 2 

Import the file `tute1.csv` from the folder data and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
tute1 <- read.csv("data/tute1.csv", header = TRUE)
View(tute1)
```

We convert the data to time series. The `[,-1]` removes the first column which contains the quarters as we don't need them now.

```{r, echo=TRUE, cache=TRUE}
mytimeseries <- ts(tute1[, -1], start = 1981, frequency = 4)
```

Construct time series plots of each of the three series and use 'facets = TRUE' .

```{r, echo=TRUE, cache=TRUE, facets = TRUE}
# to do
autoplot(mytimeseries, facets = TRUE)
```

Check what happens when you don't include `facets=TRUE`.

```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
# to do
autoplot(mytimeseries)
```

# Exercise 3

Import some monthly Australian retail data from the foler data. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
retaildata <- readxl::read_excel("data/retail.xlsx", skip = 1)
View(retaildata)
```

The second argument (`skip=1`) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

```{r, echo=TRUE, cache=TRUE}
myts <- ts(retaildata[, "A3349335T"], frequency = 12, start = c(1982, 4))
```

Explore your chosen retail time series using the following functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts)
ggAcf(myts)
```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

Your answer:
Seasonality: Der Monat Dezember hat immer jedes Jahr einen höhreren Ausschlag (warsch. Weihnachtsgeschäft).
Cyclic: -
Trend: Die Absatzzahlen steigen über die Jahre an.


# Exercise 4
Create time plots of the following time series: goog, writing, fancy, a10, h02.

    + Use `help()` to find out about the data in each series.
    + For the `goog` plot, modify the axis labels and title.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# to do
help(goog)
help(writing)
help(fancy)
help(a10)
help(h02)

autoplot(goog, xlab = "Consecutive trading days", ylab = "Stockprice", main = "Daily closing stock prices")
autoplot(writing)
autoplot(fancy)
autoplot(a10)
autoplot(h02)
```

Use the `ggseasonplot` and `ggsubseriesplot` functions to explore the seasonal patterns in the following time series:  `writing`, `fancy`, `a10`, `h02`.

- What can you say about the seasonal patterns?
- Can you identify any unusual years?


```{r, echo=TRUE}
# writing
# to do
help(writing)
ggseasonplot(writing)
ggsubseriesplot(writing)
```

Your answer:
Immer im August bricht der Absatz der Schreibwaren start ein -> Sommerferien


```{r, echo=TRUE}
# fancy
# to do
help(fancy)
ggseasonplot(fancy)
ggsubseriesplot(fancy)
```

Your answer:
Im Dezember nehmen die Absatzzahlen der Souvenire stark zu. Im trend nehmen die Absatzzahlen jedes Jahr zu. Warscheinlich Ferienzeit über die Feiertage.


```{r, echo=TRUE}
# a10
# to do
help(a10)
ggseasonplot(a10)
ggsubseriesplot(a10)
```

Your answer:
Im Monat Januar werden deutlich mehr Diabetiker Medikamente subventioniert. Warscheinlich nedingt durch die Feiertage (Weihnachten, Silvester) im Monat davor


```{r, echo=TRUE}
# h02
# to do
help(h02)
ggseasonplot(h02)
ggsubseriesplot(h02)
```

Your answer:
Nimmt stetig von Februar bis im folgejahr Januar zu und fällt dann wieder stark ab. Jährliche wiederholung.


# Exercise 5
Use the following graphics functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

    + Can you spot any seasonality, cyclicity and trend?
    + What do you learn about the series?

Example for hsales:

```{r, echo=TRUE, cache=TRUE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

  + Seasonality evident in all plots
  + Cyclicity seen in first two plots
  + No trend
  + ACF only shows seasonality. Cycle length too long to show up here.

```{r, echo=TRUE, cache=TRUE}
# to do
#usdeaths
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```

Your answer:
+Seasonality: ja erkennbar
+Cyclicity: nein
+Trend: nein
+Learned about Series: jährlicher wiederholender Zyklus mit Peak im Juli und low im Februar


```{r, echo=TRUE, cache=TRUE}
# to do
#bricksq
help(bricksq)
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq, lags = 20)
ggAcf(bricksq)
```

Your answer:
+Seasonality: ja
+Cyclicity: ja
+Trend: ja
+Learned about Data: In den Daten erkennt man im Seasonal, subseries und Acf plot eine Seasonality -> Zuhname Q1 - Q3, danach Abnahme. Der Tend nimmt allgmein zu, wird aber in den Jahren 75, 83 und 91 von einer Cyclicity unterbrochen

```{r, echo=TRUE, cache=TRUE}
# to do
#sunspotarea
help(sunspotarea)
autoplot(sunspotarea)
#ggseasonplot(sunspotarea)
#ggsubseriesplot(sunspotarea)
gglagplot(sunspotarea, , lags = 20)
ggAcf(sunspotarea, , lag.max = 70)
```

Your answer:
+Seasonality: ja
+Cyclicity:nein
+Trend: nein
+Learned about Data: Trotz des größeren Zeitraumes von 2 Jahren kann hier von einer "Seasonality" gesprochen werden, da der Zyklus konstant (10,5) Jahre ist. Besonders gut erkennbar im Lagplot. Trend steigend bis 1960 danach sinkend

```{r, echo=TRUE, cache=TRUE}
# to do
#gasoline
autoplot(gasoline)
ggseasonplot(gasoline)
#ggsubseriesplot(gasoline)
gglagplot(gasoline)
ggAcf(gasoline, lag.max = 209)
```

Your answer:
+Seasonality: ja
+Cyclicity: ja
+Trend: ja
+Learned about Data: Im Acf erkennt man deutlich die Seasonality. 6 Monate fallend und 6 monate Steigend. Im autoplot erkennt man sowohl den Steigenden Trend als auch eine Cyclicity, steigend 1991 bis 2008, dann fallend von 2008 bis 2013 und dann wieder steigend. Die fallenden Werte von 2008 bis 2013 sind warscheinlich bedingt durch die Wirtschaftskriese.

# Exercise 6

The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots.

```{r, echo=TRUE, cache=TRUE}
# White noise
set.seed(1711)
y <- ts(rnorm(50))
autoplot(y) + ggtitle("White noise")
```

```{r}
ggAcf(y)
```


```{r, echo=TRUE, cache=TRUE}
# todo
mypigs <- window(pigs, start=1990)
```
```{r}
autoplot(mypigs)
ggAcf(mypigs)
```

Your answer:
Die Werte der Lags 1-3 überschreiten die Grenze und Zeigen Autocorrelation die nichtmehr durch White Noice erzeugt werden kann.
Lag 12 und 16 sind stärkere ausschläge und könne Anzeichen für eine Seasonalität sein.


# Exercise 7

`dj` contains 292 consecutive trading days of the Dow Jones Index. Use `ddj <- diff(dj)` to compute the daily changes in the index. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?

```{r}
# todo
help(dj)
autoplot(dj)
```
```{r}
ddj <- diff(dj)
autoplot(ddj)
ggAcf(ddj, lag.max = 250)
```

Your answer:
Der Acf Plot ähnelt schon sehr stark dem von White Noice, dennoch wird die Linie an einigen Stellen überschritten.
