---
title: "Homework 1 - Time Series Toolbox"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2)
```

# Exercise 1

For the following series, find an appropriate Box-Cox (?BoxCox) transformation in order to stabilize the variance. First plot the time series and decide which of them need a transformation. 

  * `usnetelec`
  * `usgdp`
  * `mcopper`
  * `enplanements`

```{r}
# todo
autoplot(usnetelec)
lambda = BoxCox.lambda(usnetelec)
autoplot(BoxCox(usnetelec, lambda))

autoplot(usgdp)
lambda = BoxCox.lambda(usgdp)
autoplot(BoxCox(usgdp, lambda))


autoplot(mcopper)
lambda = BoxCox.lambda(mcopper)
autoplot(BoxCox(mcopper, lambda))

autoplot(enplanements)
lambda = BoxCox.lambda(enplanements)
autoplot(BoxCox(enplanements, lambda))
```


# Exercise 2

Why is a Box-Cox transformation unhelpful for the `cangas` data?

```{r}
# todo
autoplot(cangas)
```
Your answer:
Nein, eine Box-Cox transformation ist für den Fall gedacht, dass ide Varianz über die Zeit fällt oder sinkt.
Bei 'cangas' hat die Varianz aber einen 'Bauch' d.h. am Anfang ist sie gut, wird dann Schlechter und am Ende wieder besser, also kein kostanter Trend.


# Exercise 3

What Box-Cox transformation would you select for **your** retail data from Exercise 4 (3?) in Visualization?

```{r}
# Example 
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))
autoplot(myts)
myts %>% BoxCox(lambda = 0) %>% autoplot()
```

From visual inspection, a log transformation would be appropriate here. It also makes sense, as retail expenditure is likely to increase proportionally to population, and therefore the seasonal fluctuations are likely to be proportional to the level of the series. It has the added advantage of being easier to explain than some other transformations. Finally, it is relatively close to the automatically selected value of `BoxCox.lambda(myts)` $= `r round(BoxCox.lambda(myts),3)`$.

If you have selected a different series from the retail data set, you might choose a different transformation.

Your answer:
```{r}
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[, "A3349335T"], frequency = 12, start = c(1982, 4))

autoplot(myts)

lambda = BoxCox.lambda(myts)
autoplot(BoxCox(myts, lambda))

print(lambda)
```
Im Grunde lassen sich die eigenen Daten gleich erklären wie die aus dem Beispiel. Der Wert steigt mit der Zahl der Bevölkerung an. Dementsprechend fallen auch die Sesonalen Peaks bei höheren werten proportional höher aus. Der log gleicht ides am besten aus. Der autowert von lambda liegt auch bei 0,19 was am ehsten einem log entspricht.


# Exercise 4

Calculate the residuals (?residuals) from a seasonal naive forecast applied to the quarterly Australian beer production data from 1992. The following code will help.

```{r}
# todo
help(ausbeer)
autoplot(ausbeer)

#beer <- as_tsibble(beer)

lbd = BoxCox.lambda(myts)
fit <- snaive(ausbeer, lambda=lbd, h=40)
fit2 <- snaive(ausbeer, lambda=0, h=40)
fit3 <- snaive(ausbeer, h=40)
autoplot(fit)
autoplot(fit2)
autoplot(fit3)
```

Test if the residuals are white noise and normally distributed.

```{r}
checkresiduals(fit)
```

What do you conclude?

Your answer:
Der ACF Lag-Plot zeigt, dass eine Korrelation vorhanden ist, da der positive Grenzwert mehrfach überschritten wird. Desshalb handelt es sich nicht um weißes Rauschen, was bedeutet es sind noch Informationen in den Daten die nicht für die Vorhersage benutzt werden.


# Exercise 5

Are the following statements true or false? Explain your answer.

> a. Good forecast methods should have normally distributed residuals.

Your answer: Ja, eine gute Vorhersage sollte, wie auch die Daten, eine normalveteilte Fhelerverteilung haben, da diese auf dem Zufall baisert.

> b. A model with small residuals will give good forecasts.

Your answer: Richtig, kleine Residuals bedeuten, dass die Zusammenhäbge in den Daten erkann wurden und daher auch gute Vorhersagen getroffen werden können

> c. The best measure of forecast accuracy is MAPE.

Your answer: Falsch, Mape wird z.B. durch Ausreiser stark beeinflusst. Eine Methodik die dies berücksichtigt durch normalisierung der Daten o.Ä. wäre besser geeignet. Ein Beispiel dafür wäre RSME, also die Mittlere Quadratische Abweichung.

> d. If your model doesn't forecast well, you should make it more complicated.

Your answer: Falsch. Ein komplexeres Modell bedeutet nicht direkt eine bessere Vorhersage. Man sollte viel mehr den Grund für die Schlechte Vorhersage finden z.b. durch bessere Vorverarbeitung der Daten.

> e. Always choose the model with the best forecast accuracy as measured on the test set.

Your answer: Richtig, insofern der Testdatensatz umfangreich genug ist und die gleichen Eigenschaften besitzt wie die zu erwartenden Daten.  

# Exercise 6

For your retail time series (from Exercise 4):

Split the data into two parts using

```{r}
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349335T"], frequency=12, start=c(1982,4))
myts.train <- window(myts, end=c(2010,12))
myts.test <- window(myts, start=2011)
```

Check that your data have been split appropriately by producing the following plot.

```{r}
autoplot(myts) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")
```

Calculate forecasts using `snaive` applied to `myts.train`.

```{r}
# todo
forecast = snaive(myts.train, h=36)
autoplot(myts) +
  autolayer(myts.train, series="train") +
  autolayer(forecast, series="forecast") +
  autolayer(myts.test, series="test") 

```

Compare the accuracy of your forecasts against the actual values stored in `myts.test`.
(?accuracy) 

```{r}
# todo

accuracy(forecast, myts.test)


```

The number to look at here is the test set RMSE of 71.443. That provides a benchmark for comparison when we try other models.

Check the residuals. Do the residuals appear to be uncorrelated and normally distributed?

```{r}
# todo
 checkresiduals(forecast)
```

Your answer:
Nein, die Korrelation überschreitet konstant den Grenzwert, vor alle für Werte mit einem Lag<10.
Die Residuals sind auch nicht Normalverteilt.

How sensitive are the accuracy measures to the training/test split?


# Exercise 7

`visnights` contains quarterly visitor nights (in millions) from 1998-2015 for eight regions of Australia.

Use `window()` to create three training sets for `visnights[,"QLDMetro"],` omitting the last 1, 2 and 3 years; call these train1, train2, and train3, respectively.

```{r}
train1 <- window(visnights[, "QLDMetro"], end = c(2015, 3))
train2 <- window(visnights[, "QLDMetro"], end = c(2015, 2))
train3 <- window(visnights[, "QLDMetro"], end = c(2015, 1))
```

Compute one year of forecasts for each training set using the `snaive()` method. Call these `fc1`, `fc2` and `fc3`, respectively.

```{r}
# todo

fc1 = snaive(train1)
fc2 = snaive(train2)
fc3 = snaive(train3)
 
```

Use `accuracy()` to compare the MAPE over the three test sets. Comment on these.
First we will copy the actual data into a variable. Then we can do an accuracy comparison.

```{r}
qld <- visnights[, "QLDMetro"]
# todo
accuracy(fc1, qld)
accuracy(fc2, qld)
accuracy(fc3, qld) 
```

This should give similar results to this consolidated results table.

```
                     ME     RMSE      MAE        MPE     MAPE      MASE       ACF1
Training set  0.1637836 1.742687 1.360271  0.4384347 7.357322 1.0000000 0.06643175
Test set fc1 -1.3010774 1.301077 1.301077 -6.9956861 6.995686 0.9564839         NA
Test set fc2 0.08383478 1.387447 1.384912 -0.4063445 6.589342 1.019346 -0.50000000
Test set fc3 0.06202858 1.132896 0.9294135 -0.237857 4.425934 0.6738562 -0.51548610
```

The lower MAPE value for "fc3" indicates a better result when we use the previous 3 values for the `snaive()` prediction.

# Exercise 8

Use the Dow Jones index (data set `dowjones`) to do the following:

Produce a time plot of the series.

```{r}
# todo
help(downjones)
autoplot(dowjones)
```

Produce forecasts using the drift method and plot them.

Let's assume we want to forecast the next 5, 10 and 15 values.

```{r}
# todo

dowfc1 <- rwf(dowjones, drift=TRUE, h=5)
dowfc2 <- rwf(dowjones, drift=TRUE, h=10) 
dowfc3 <- rwf(dowjones, drift=TRUE, h=15) 

```

Then we can plot these values.

```{r eval = FALSE}
# remove eval=FALSE
autoplot(dowjones) +
  autolayer(dowfc1, PI=FALSE, series="Drift 5") +
  autolayer(dowfc2, PI=FALSE, series="Drift 10") +
  autolayer(dowfc3, PI=FALSE, series="Drift 15") +
  xlab("Time") + ylab("Closing Price (US$)") +
  ggtitle("Dow Jones index") +
  guides(colour=guide_legend(title="Forecast"))
```

We show that the forecasts are identical to extending the line drawn between the first and last observations.

We can plot the forecasts in a different order, so the shorter forecasts are superimposed, showing the lines are the same.

```{r eval = FALSE}
# remove eval=FALSE
autoplot(dowjones) +
  autolayer(dowfc3, PI=FALSE, series="Drift 15") +
  autolayer(dowfc2, PI=FALSE, series="Drift 10") +
  autolayer(dowfc1, PI=FALSE, series="Drift 5") +
  xlab("Time") + ylab("Closing Price (US$)") +
  ggtitle("Dow Jones index") +
  guides(colour=guide_legend(title="Forecast"))
```

Try using some of the other benchmark functions to forecast the same data set. Which do you think is best? Why?

The time series isn't seasonal, so the seasonal naive method is not viable. However, we can use the mean and naive methods.

```{r}
# todo

dowfc_naive = naive(dowjones) 
dowfc_meanf = meanf(dowjones)
 
autoplot(dowfc_naive)
autoplot(dowfc_meanf)

accuracy(dowfc_naive)
accuracy(dowfc_meanf)

```

The three values will be very different here. The Mean will use the data set, so is unlikely to follow the current trendline.

Drift ist wohl am besten, da es den Trend der Vorherigen Daten fortsetzt. Mean ist nicht wirklich aussagekräftig als Vorhersage, da es nur den Wert mittelt und die Abweichungen sehr groß sind.

# Exercise 9

Consider the daily closing IBM stock prices (data set `ibmclose`).

Produce some plots of the data in order to become familiar with it.

```{r}
# todo
help(ibmclose)
print(ibmclose)
```
```{r}
autoplot(ibmclose)
gglagplot(ibmclose)
ggAcf(ibmclose)
```

Split the data into a training set of 300 observations and a test set of 69 observations.

```{r}
# todo
print(length(ibmclose))
ibm.train <- window(ibmclose, end = 300)
ibm.test <- window(ibmclose, start = 301)

```

Try using various benchmark methods to forecast the training set and compare the results on the test set. Which method did best?

```{r eval = FALSE}
# remove eval=FALSE
h <- length(ibm.test)
m.f <- meanf(ibm.train, h=h)
# todo

rw.f <- naive(ibm.train, h=h)
rwd.f <- rwf(ibm.train, drift = TRUE, h=h)
 


autoplot(ibmclose) +
  xlab("Day") +
  ggtitle("Daily closing IBM stock prices") +
  autolayer(m.f$mean, col=2, series="Mean method") +
  autolayer(rw.f$mean, col=3, series="Naive method") +
  autolayer(rwd.f$mean, col=4, series="Drift method")

# accuracy(m.f,ibm.test)
# todo
 accuracy(m.f,ibm.test)
 accuracy(rw.f,ibm.test)
 accuracy(rwd.f,ibm.test)
```


Check the residuals of your preferred method. Do they resemble white noise?

```{r}
# todo
checkresiduals(rwd.f) 
```
Ein ein paar wenigen Stellen überschreitet die Korrelation zwar noch die Schwelle, ist aber größtenteils als White Noise erkennbar. Auch die Residuals sind nah an einer Normalverteilung.



# Exercise 10

Consider the sales of new one-family houses in the USA, Jan 1973 -- Nov 1995 (data set `hsales`).

Produce some plots of the data in order to become familiar with it.

```{r}
# todo
help(hsales)
autoplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
```

Split the `hsales` data set into a training set and a test set, where the test set is the last two years of data.

```{r}
# todo
print(length(hsales))
print(hsales)
hs.train <- window(hsales, start = 1993)
hs.test <- window(hsales, end = 1994)
```

Try using various benchmark methods to forecast the training set and compare the results on the test set. Which method did best?

```{r}
# todo
h <- length(hs.train)
s.f <- snaive(hs.train, h=h)
m.f <- meanf(hs.train, h=h)
rw.f <- naive(hs.train, h=h)
rwd.f <- rwf(hs.train, drift = TRUE, h=h)


autoplot(hsales) +
  xlab("Month") +
  ggtitle("Monthly sales of one-family houses") +
  autolayer(s.f$mean, series="SNaive method") +
  autolayer(m.f$mean, series="Mean method") +
  autolayer(rw.f$mean, series="Naive method") +
  autolayer(rwd.f$mean, series="Drift method")
```
SNaive Methode liefert das Beste Ergebnis.

In terms of accuracy measures on the test set, the seasonal naive method does better.

Check the residuals of your preferred method. Do they resemble white noise?

```{r eval = FALSE}
# remove eval=FALSE
checkresiduals(s.f)
```
Kein White Noice. Zwar wird die Schwelle bei der Korrelatio nicht überschritten, allerdings kann man einen Trend erkennen, der von anfangs sehr positiv ins negative läuft..

